--
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Here is my submission for the reproducible research peer assessment 1

```{r packages, echo = TRUE,results='hide'}
library(lubridate)
library(dplyr)
library(ggplot2)
```


## Loading and preprocessing the data

```{r load, echo = TRUE}
activity <- read.csv("activity.csv", na.strings="NA")
activity$date <- as.Date(activity$date)
```

## What is mean total number of steps taken per day?
```{r, echo = TRUE}
steps.sum <- aggregate(x = activity["steps"],
                     FUN = sum,
                     by = list(Group.date = activity$date))
hist(steps.sum$steps,main="Total Number of Steps per Day Histogram",xlab="Total Number of Steps per Day")
```


```{r, echo = TRUE}
mean <- mean(steps.sum$steps, na.rm=TRUE)
```

The mean number of steps is `r mean`.
```{r, echo = TRUE}
median <-median(steps.sum$steps, na.rm=TRUE)
```
The median number of steps is `r median`.

## What is the average daily activity pattern?

```{r, echo = TRUE}
steps.interval <- aggregate(x = activity["steps"],
                     FUN = mean,
                     by = list(Group.interval = activity$interval),
                     na.rm=TRUE)
plot(steps.interval$Group.interval, steps.interval$steps, type = "l", xlab="Interval", ylab="Average Steps")
abline(v=steps.interval$Group.interval[which.max(steps.interval$steps)], col="red")
```

```{r, echo = TRUE}
interval<-steps.interval$Group.interval[which.max(steps.interval$steps)]
```

The 5-minute interval with the maximum number of steps is `r interval`.

## Imputing missing values

```{r, echo = TRUE}
numberna <- sum(is.na(activity))
```

The total number of NA is `r numberna`.

```{r}
new <- activity
for (i in steps.interval$Group.interval) {
  new[new$interval == i & is.na(new$steps), ]$steps <- steps.interval$steps[steps.interval$Group.interval == i]
}
sum(is.na(new))
```


```{r, echo = TRUE}
newsteps.sum <- aggregate(x = new["steps"],
                     FUN = sum,
                     by = list(Group.date = new$date))
```


```{r, echo = TRUE}
hist(newsteps.sum$steps,main="Total Number of Steps per Day Histogram w/o NA",xlab="Total Number of Steps per Day")
```


```{r, echo = TRUE}
newmean <- mean(newsteps.sum$steps, na.rm=TRUE)
```

The mean number of steps is `r newmean` when calculated without missing values.
```{r, echo = TRUE}
newmedian <-median(newsteps.sum$steps, na.rm=TRUE)
```
The median number of steps is `r newmedian` when calculated without missing values.

Imputing missing data increases the total number of steps but does not impact the mean or the median. 

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo = TRUE}
new$week <- ifelse(weekdays(new$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
```

```{r, echo = TRUE}
new.interval <- aggregate(x = new["steps"],
                     FUN = mean,
                     by = list(Group.interval = new$interval,Group.week=new$week),
                    )
new.interval$Group.week<-factor(new.interval$Group.week)
ggplot(data = new.interval)+
  aes(x=Group.interval,y= steps, group = Group.week, col = Group.week)+
  geom_line()+
  facet_grid(Group.week~.)+
  labs(title= "Average Steps Per Interval by Days",x="Interval", y="Average Steps") 
```